Python as a machine learning language: experiences from the scikit-learn project

### 2013-06-16 paper.tex
Move duck typing and trim / merge "Cython for inner loops".
@glouppe: I moved the duck typing paragraph to the meta-estimator
subsection, where I think the concept could introduced nicely (e.g.,
by saying that duck typing allows to use any estimator as base
estimator).

@larsmans: I trimmed and moved the part "Cython for inner loops".
BTW, SGD uses loss functions implemented in Cython but they are passed
as argument from the pure Python side. Yet, the performance seems
reasonable. I wonder if we could use a similar approach for kernel-based
methods...

@@

\section{Implementation}

\label{sec:implementation}

@@

% scikit-learn is primarily implemented in Python and Cython
% ,
% a language that extends Python with static typing
% and a compiler that produces C extension modules
% for the Python runtime system.
% In addition, it includes (modified versions of)
% the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
% used for training support vector machines
% and logistic regression models \citep{chang2011libsvm, fan2008}.
% These libraries are both written in C{}\verb!++!
% and wrapped using Cython modules.

% As much as possible is implemented in ``pure'' Python
% with NumPy for numerical work,
% to avoid the compile-edit-debug cycle needed for Cython
% and to keep the codebase readable for as large an audience as possible.
% However, not all algorithms are easily and efficiently expressed
% as NumPy vector operations
% (among other factors, because they often have linear space requirements).
% Algorithms that fall into this category have to be implemented in Cython
% to obtain decent performance and scalability.
% Examples include $k$-means, stochastic gradient descent for linear models
% and some graph-based clustering algorithms.

% While object-oriented techniques and design patterns
% are used throughout scikit-learn,
% they mostly serve to facilitate code reuse and are not considered a project goal.
% Where possible, the Python principle of ``duck typing'' is exploited
% to simplify the implementation and skip the introduction of superfluous classes.
% This allows for extensibility and flexibility at the same time:
% user-defined estimators that follow scikit-learn conventions
% should be usable in pipelines and other composite objects
% without them actually inheriting from scikit-learn base classes.

@@

\subsection{Cython for inner loops}

A problem arising from the use of Cython
concerns the mismatch between on the one hand Python functions,
which are first-class runtime objects endowed with methods and metadata
that identify them as ``callable'',
that dynamically check their number and types of arguments,
and that can return any Python type of object;
and compiled C or Cython functions on the other hand,
which exhibit static typing.
Where it may sometimes be convenient for estimators
to take functions as arguments
(e.g., in the case of kernel machine learning,
where one might like to use a custom kernel),
but such usage is not at present possible and may never be.
Passing C or Cython-compiled functions
would require an API for the kernel method
to be implemented at the C or Cython level,
only to be used from compiled code.
Passing arbitrary Python functions would be possible,
but performance can be expected to degrade significantly
because of the runtime checks needed for each invocation of the kernel function.
(The current solution in scikit-learn is to let the user pass
not a kernel function,
but a precomputed Gram matrix for an entire set of objects.
The linear, polynomial and RBF kernels are treated as special cases in
the \textsf{LIBSVM}-based SVM learner,
where they are implemented in C{}\verb!++!.)

>>>>

\textcolor{red}{Point out that any external estimator can be used as base
estimator thanks to duck typing}\\

@@

\section{Implementation}
\label{sec:implementation}

\subsection{Cython for performance}

@@

A problem arising from the use of Cython, however, is that passing pure Python
functions as argument, while possible, degrades performance significantly, due
to the runtime checks needed for each function invocation. This especially
problematic, for example for kernel methods, which would benefit from accepting
a kernel function as argument. A potential solution to this problem would be for
\sklearn to provide a Cython-level API, thereby alllowing the user to pass
Cython-implemented kernel functions.

@@
